---
title: "Lab 04"
output: github_document
always_allow_html: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1.Read the data

```{r packages}
library(data.table)
library(tidyverse)
```


```{r, cache=TRUE}
if (!file.exists("met_all.gz"))
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz",
    destfile = "met_all.gz",
    method   = "libcurl",
    timeout  = 60
    )
met <- data.table::fread("met_all.gz")
```

## 2.Prepare the data

```{r data-prep, cache=TRUE}
#Remove temperatures less than -17C
met <- met[temp>=-17]

#Make sure there are no missing data in the key variables coded as 9999, 999, etc
met[,range(temp)]
met[,range(rh, na.rm = TRUE)]
met[,range(wind.sp, na.rm = TRUE)]
met[,range(vis.dist, na.rm = TRUE)]
met[,range(dew.point, na.rm = TRUE)]
met[,range(lat, na.rm = TRUE)]
met[,range(lon, na.rm = TRUE)]
met[,range(elev, na.rm = TRUE)]

met[elev==9999.0, elev:=NA]

#Generate a date variable using the functions as.Date() (hint: You will need the following to create a date paste(year, month, day, sep = "-")).
met[, ymd :=as.Date(paste(year,month,day,sep="-"))]

#Using the data.table::week function, keep the observations of the first week of the month.
met[, table(week(ymd))]
met <-met[week (ymd) ==31]

#Compute the mean by station of the variables temp, rh, wind.sp, vis.dist, dew.point, lat, lon, and elev.
met_avg <-met[, .(
  temp = mean(temp,na.rm=TRUE),
  rh = mean(rh,na.rm=TRUE),
  wind.sp = mean(wind.sp,na.rm=TRUE),
  vis.dist = mean(vis.dist,na.rm=TRUE),
  dew.point = mean(dew.point,na.rm=TRUE),
  lat = mean(lat,na.rm=TRUE),
  lon = mean(lon,na.rm=TRUE),
  elev = mean(elev,na.rm=TRUE), USAFID
), by = "USAFID"]

#Create a region variable for NW, SW, NE, SE based on lon = -98.00 and lat = 39.71 degrees
met_avg[lat >= 39.71 & lon <=-98.00, region := "Northwest"]
met_avg[lat < 39.71 & lon <=-98.00, region := "Southwest"]
met_avg[lat >= 39.71 & lon >-98.00, region := "Notrheast"]
met_avg[lat < 39.71 & lon >-98.00, region := "Southeast"]

met_avg[,table(region,useNA ="always")]

#Create a categorical variable for elevation as in the lecture slides
met_avg[, elev_cat := fifelse(elev > 252, "high", "low")]

#Deleting extra column
met_avg <- met_avg[,-1]

```

## 3.Use geom_violin to examine the wind speed and dew point temperature by region

```{r violin-temp}
ggplot(met_avg,mapping = aes(x =1, y = wind.sp)) +
  geom_violin() +
  facet_grid(~region)
```

## 4. Use geom_jitter with stat_smooth to examine the association between dew point temperature and wind speed by region

```{r}
ggplot()+
  geom_jitter(met_avg[!is.na(wind.sp) & !is.na(dew.point)],
       mapping = aes(x=dew.point, y=wind.sp,color=region))+
  stat_smooth(met_avg,mapping=aes(x=dew.point,y=wind.sp,linetype=region))
```


## 5. Use geom_bar to create barplots of the weather stations by elevation category coloured by region

```{r}
ggplot()+
  geom_bar(met_avg[!is.na(elev_cat)],
           mapping = aes(x=elev_cat,fill=region),position="dodge")+
  labs(
    x="elevation categories",
    y="Frequency",
    title ="bar chart of elev_cat colored by region"
  )+
  scale_fill_brewer(palette="Accent")
```


## 6. Use stat_summary to examine mean dew point and wind speed by region with standard deviation error bars
```{r}
ggplot(met_avg[!is.na(wind.sp) & !is.na(dew.point)],
       mapping = aes(x = region, y = wind.sp)) +
  stat_summary(fun.data = "mean_sdl") +
  stat_summary(fun.data ="mean_sdl",geom = "errorbar")
```

## 7. Make a map showing the spatial trend in relative h in the US
```{r}
library(leaflet)

temp.pal <- colorNumeric(c('darkgreen','goldenrod','brown'), domain=met_avg$rh)

leaflet(met_avg) %>%
  addProviderTiles('CartoDB.Positron') %>%
  addCircles(
    lat = ~lat, lng=~lon,
    label = ~paste0(round(rh,2), ' rh'), color = ~ temp.pal(rh),
    opacity = 1, fillOpacity = 1, radius = 500
    ) %>%
  addLegend('bottomleft', pal=temp.pal, values=met_avg$rh,
          title='Temperature, C', opacity=1)


```

## 8.Use a ggplot extension

```{r, cache=TRUE}
library(gganimate)

ggplot()+
  geom_point(met[!is.na(rh) & !is.na(temp)],mapping=aes(rh,temp))+
  labs(x="reletive height",y="temperature")+
  transition_time(hour)
```

