---
title: "Lab05"
author: Jiayi Nie
output: github_document
always_allow_html: TRUE
---

```{r package}
library(tidyverse)
library(data.table)
library(leaflet)

```

```{r read the data, cache= TRUE}
met<- data.table::fread("../Lab04/met_all.gz")

# Download the data
stations <- fread("ftp://ftp.ncdc.noaa.gov/pub/data/noaa/isd-history.csv")
stations[, USAF := as.integer(USAF)]

# Dealing with NAs and 999999
stations[, USAF   := fifelse(USAF == 999999, NA_integer_, USAF)]
stations[, CTRY   := fifelse(CTRY == "", NA_character_, CTRY)]
stations[, STATE  := fifelse(STATE == "", NA_character_, STATE)]

# Selecting the three relevant columns, and keeping unique records
stations <- unique(stations[, list(USAF, CTRY, STATE)])

# Dropping NAs
stations <- stations[!is.na(USAF)]

# Removing duplicates
stations[, n := 1:.N, by = .(USAF)]
stations <- stations[n == 1,][, n := NULL]
```


merging the data

```{r}
met <- merge(
  # Data
  x     = met,      
  y     = stations, 
  # List of variables to match
  by.x  = "USAFID",
  by.y  = "USAF", 
  # Which obs to keep?
  all.x = TRUE,      
  all.y = FALSE
  )
```

## Question1: Representative station for the US

What is the median station in terms of temperature, wind speed, and atmospheric pressure? Look for the three weather stations that best represent continental US using the quantile() function. Do these three coincide?

First, generate a representative version of each station. We will use the averages (median could also be a good way to represent it, but it will depend on the case).

```{r collapsing-by-station}
station_averages <- met[,.(
  temp      = mean(temp, na.rm = TRUE),
  wind.sp   = mean(wind.sp, na.rm = TRUE),
  atm.press = mean(atm.press, na.rm = TRUE),
  lon = mean(lon, na.rm = TRUE),
  lat = mean(lat, na.rm = TRUE)
), by = .(USAFID)]
```

Now, we need to identify the median per variable

```{r quantile}
medians <- station_averages[,.(
  temp_50      = quantile(temp, probs = .5, na.rm = TRUE),
  wind.sp_50   = quantile(wind.sp, probs = .5, na.rm = TRUE),
  atm.press_50 = quantile(atm.press, probs = .5, na.rm = TRUE)
)]

```

Now we can find the stations that are the closest to these. (hint: which.min())

```{r}
station_averages[, temp_dist := abs(temp - medians$temp_50)]
median_temp_station <- station_averages[order(temp_dist)][1]
median_temp_station

station_averages[, wind.sp_dist := abs(wind.sp - medians$wind.sp_50)]
median_wind.sp_station <- station_averages[order(wind.sp_dist)][1]
median_wind.sp_station

station_averages[, atm.press_dist := abs(temp - medians$atm.press_50)]
median_atm.press_station <- station_averages[order(atm.press_dist)][1]
median_atm.press_station
```

720458,720929,722238 respectively so these stations are not coincide.

## Question2: Representative station per state

We first need to recover the state variable, by merging

```{r}
station_averages <- merge(
  x = station_averages, y = stations,
  by.x = "USAFID", by.y = "USAF",
  all.x = TRUE, all.y = FALSE
  )
```

Now we can compute the median per state

```{r}
station_averages[, temp_50 := quantile(temp, probs = .5, na.rm = TRUE), by = STATE]
station_averages[, wind.sp_50 := quantile(wind.sp, probs = .5, na.rm = TRUE), by = STATE]
station_averages[, atm.press_50 := quantile(atm.press, probs = .5, na.rm = TRUE), by = STATE]


```

Now, the euclidean distance… $\sqrt{\sum_i(x_i - y_i)^2}$

```{r}
station_averages[, eudist := sqrt(
  (temp - temp_50)^2 + (wind.sp - wind.sp_50)^2 + (atm.press - atm.press_50)^2
  )]
station_averages_state <- station_averages[ , .SD[which.min(eudist)], by = STATE]

station_averages_state

```


## Question 3: In the middle?

```{r}
#calculate median for lon and lat per state
station_averages[, lon_50 := quantile(lon, probs = .5, na.rm = TRUE), by = STATE]
station_averages[, lat_50 := quantile(lat, probs = .5, na.rm = TRUE), by = STATE]

#now calculate the distance
station_averages[, latlon_dis := sqrt(
  (lat-lat_50)^2 + (lon - lon_50)^2
)]

midpoint<- station_averages[, .SD[which.min(latlon_dis)],by =STATE]

midpoint

#combine the dataset

c_state <- rbind(station_averages_state, midpoint, fill = TRUE)

```

```{r}
library(leaflet)

leaflet(c_state) %>%
  addProviderTiles('CartoDB.Positron') %>%
  addCircles(
    lat = ~lat, lng = ~lon, 
    color = "lightgreen",
    opacity = 1,
    fillOpacity = 1,
    radius=100
)
```

## Question4: Means of means

Going back to the met dataset.

```{r}
met[, state_temp := mean(temp, na.rm = TRUE), by = STATE]
met[, state_wind.sp := mean(wind.sp, na.rm = TRUE), by = STATE]
met[, state_atm.press := mean(atm.press, na.rm = TRUE), by = STATE]

met[, temp_cat := fifelse(
  state_temp < 20, "low-temp", 
  fifelse(state_temp < 25, "mid-temp", "high-temp"))
  ]

```

Let’s make sure that we don’t have NAs

```{r}
table(met$temp_cat, useNA = "always")

```

Now, let’s summarize

```{r}
tab <- met[, .(
  N_entries  = .N,
  N_stations = length(unique(USAFID)),
  avg_temp = mean(state_temp),
  avg_wind.sp = mean(state_wind.sp),
  avg_atm.press = mean(state_atm.press)
), by = temp_cat]

knitr::kable(tab)

```




