---
title: "Lab06"
author: Jiayi Nie
output: github_document
always_allow_html: TRUE
---

```{r set-up}
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(repos = c(CRAN = "http://cran.rstudio.com"))

```


```{r packages, include=FALSE}
library(dplyr)
library(ggplot2)
library(tidytext)
library(tidyverse)
library(forcats)
```

## Read the data

```{r get data}
fn <- "mtsamples.csv"

if(!file.exists(fn))
  download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/00_mtsamples/mtsamples.csv", destfile = fn)

mtsamples <- read.csv(fn)
mtsamples <- as_tibble(mtsamples)
```

## Q1:How many and how are specialities distributed?

```{r dis-of-specialities}
specialties <- mtsamples %>%
  count(medical_specialty)

specialties %>%
  arrange(desc(n)) %>%
  top_n(15) %>%
  knitr::kable()
  
```

There are `r nrow(specialties)` specialties. Let's take a look at the distribution.

```{r}
ggplot(mtsamples, aes(medical_specialty))+
  geom_histogram(stat="count") +
  coord_flip()
```

these are not evenlt (uniformly) distributed.

## Q2

```{r Tokenize the the words in the transcription column, cache=TRUE}
mtsamples %>%
  unnest_tokens(output = word, input = transcription) %>%
  count(word, sort = TRUE) %>%
  top_n(20)%>%
  ggplot(aes(n,fct_reorder(word,n)))+
    geom_col()
```

The word "patient" is seems to be important, but we observe a lot of stopwords.

## Q3

```{r Redo visualization but remove stopwords before}
mtsamples %>%
  unnest_tokens(output = word, input = transcription) %>%
  count(word, sort = TRUE) %>%
  anti_join(stop_words, by= "word")%>%
  top_n(20)%>%
  ggplot(aes(n,fct_reorder(word,n)))+
    geom_col()
```

looking better , but we don't like the number

```{r}

```

